### 2013-07-26 doc/taxize_withcode_plos.Rnw
edited readme, making changes in ms according to reviews from naim and karthik, still working

\section*{Introduction}
Evolution by natural selection has led to a hierarchical relationship among all living organisms.  Thus, species are categorized using a taxonomic hierarchy, starting with the binomial species name (e.g, \emph{Homo sapiens}), moving up to genus (\emph{Homo}), then family (\emph{Hominidae}), and on up to Domain (\emph{Eukarya}). Biologists, whether studying organisms at the cell, organismal, or community level, can put their study taxa into taxonomic context, allowing them to know close and distant relatives, find relevant literature, and more.

The use of taxonomic names is, unfortunately, not straightforward. Taxonomic names often vary due to name changes at the generic or specific levels, lumping or splitting lower taxa (genera, species) among higher taxa (families), and name spelling changes. For example, a study found that a compilation of 308,000 plant observations from 51 digitized herbarium records had 22,100 unique taxon names, of which only ~13,000 were accepted names \cite{weiser2007,boyle2013}. In addition, there is no one authoritative taxonomic names source for all taxa - although, there are taxon specific sources that are used by many scientists. Different sources (e.g., uBio, Tropicos, ITIS) may use different accepted names for the same taxon. For example, while the Integrated Taxonomic Information Service (ITIS) has \emph{Helianthus x glaucus} as an accepted name, The Plant List \cite{theplantlist} has that name as unresolved. But \emph{Helianthus glaucus} is an accepted name in The Plant List, while ITIS does not list this name.

@@

Taxonomic data is getting easier to obtain through the web (e.g., \cite{eol}). However, there are a number of good reasons to obtain taxonomic information programatically rather than through a web interface. First, if you have more than a few names to lookup on a website, it can take quite a long time to enter each name, get data, and repeat for each species. Programatically getting taxonomic names solves the problem by looping over a list of names. In addition, doing taxonomic searching, etc. becomes reproducible. With increasing reports of irreproducibility in science \cite{stodden2010,zimmer2012}, it is extremely important to make science workflows repeatable. Science workflows can now easily incorporate text, code, and images in a single executable document \cite{yihui2013}.

The R language is a widely used language by biologists, and now has over 4,000 packages on the Comprehensive R Archive Network (CRAN) and more than 2,500 packages on other repositories to extend R. R is great for manipulating, visualizing and fitting statistical models to data. \cite{gentleman_bioconductor:_2004} gives a detailed discussion of advantages of R in computational biology. Getting data from the web will be increasingly common as more and more data gets moved to the cloud. Therefore there is a need to get data from the web directly into R. Increasingly, data is available from the web via application programming interfaces (API). These allow computers to talk to one another using code that is not human readable, but is machine readable. Web APIs often define a number of methods that allow users to search for a species name, or retrieve the synonyms for a species name, for example. A further advantage of APIs is that they are language agnostic, meaning that data can be consumed in almost any computing context, allowing users to interact with the web API without having to know the details of the code. Moreover data can be accessed from every computer, whereas for example an Excel file can only be opened in a few programs.

>>>>

\section*{Introduction}
Evolution by natural selection has led to a hierarchical relationship among all living organisms.  Thus, species are categorized using a taxonomic hierarchy, starting with the binomial species name (e.g, \emph{Homo sapiens}), moving up to genus (\emph{Homo}), then family (\emph{Hominidae}), and on up to Domain (\emph{Eukarya}). Biologists, whether studying organisms at the cell, organismal, or community level, can put their study objects into taxonomic context, allowing them to know close and distant relatives, find relevant literature, and more.

The use of taxonomic names is, unfortunately, not straightforward. Taxonomic names often vary due to name revisions at the generic or specific levels, lumping or splitting lower taxa (genera, species) among higher taxa (families), and name spelling changes. For example, a study found that a compilation of 308,000 plant observations from 51 digitized herbarium records had 22,100 unique taxon names, of which only ~13,000 were accepted names \cite{weiser2007,boyle2013}. In addition, there is no one authoritative taxonomic names source for all taxa - although, there are taxon specific sources that are used by many scientists. Different sources (e.g., uBio, Tropicos, ITIS) may use different accepted names for the same taxon. For example, while the Integrated Taxonomic Information Service (ITIS) has \emph{Helianthus x glaucus} as an accepted name, The Plant List \cite{theplantlist} has that name as unresolved. But \emph{Helianthus glaucus} is an accepted name in The Plant List, while ITIS does not list this name.

@@

Taxonomic data is getting easier to obtain through the web (e.g., \cite{eol}). However, there are a number of good reasons to obtain taxonomic information programatically rather than through a web interface. First, if you have more than a few names to lookup on a website, it can take quite a long time to enter each name, get data, and repeat for each species. Programatically getting taxonomic names solves the problem by looping over a list of names. In addition, doing taxonomic searching, etc. becomes reproducible. With increasing reports of irreproducibility in science \cite{stodden2010,zimmer2012}, it is extremely important to make science workflows repeatable. Science workflows can now easily incorporate text, code, and images in a single executable document \cite{yihui2013}. Reproducible documents should become mainstream in biology to avoid mistakes, and make collaboration easier.

The R language is a widely used language by biologists, and now has over 5,000 packages on the Comprehensive R Archive Network (CRAN) and more than 2,500 packages on other repositories to extend R. R is great for manipulating, visualizing and fitting statistical models to data. \cite{gentleman_bioconductor:_2004} gives a detailed discussion of advantages of R in computational biology. Getting data from the web will be increasingly common as more and more data gets moved to the cloud. Therefore there is a need to get data from the web directly into R. Increasingly, data is available from the web via application programming interfaces (API). These allow computers to talk to one another using code that is not human readable, but is machine readable. Web APIs often define a number of methods that allow users to search for a species name, or retrieve the synonyms for a species name, for example. A further advantage of APIs is that they are language agnostic, meaning that data can be consumed in almost any computing context, allowing users to interact with the web API without having to know the details of the code. Moreover data can be accessed from every computer, whereas for example an Excel file can only be opened in a few programs.
